---
title: "RWorksheet_Pabriaga#4b"
author: "John Martin S. Pabriaga"
date: "2024-10-28"
output:
  pdf_document: default
---

```{r, for}
#1.
vectorA <- c(1, 2, 3, 4, 5)

zero_matrix <- matrix(0, nrow = 5, ncol = 5)

result_matrix <- matrix(0, nrow = 5, ncol = 5)

for (i in 1:5) {
  for (j in 1:5) {
    result_matrix[i, j] <- abs(vectorA[j] - i)
  }
}

result_matrix
```
```{r, loop}
#2. 
matrix_size <- 5
matrix <- matrix(0, nrow = matrix_size, ncol = matrix_size)

for (i in 1:matrix_size) {
  for (j in 1:matrix_size) {
    matrix[i, j] <- abs(i - j)
  }
}

cat("Transformed Matrix:\n")
print(matrix)
cat("\n")

rows <- 5
cat("Star Pattern:\n")

for (i in 1:rows) {
  cat(rep("*", i), sep = " ", "\n")
}

```
```{r}
first <- as.integer(readline(prompt = "Enter the first number: "))

if (is.na(first)) {
  cat("Please enter a valid integer.\n")
} else {
  cat(first, "", sep = "") 
  fibonacci <- c(first)

  repeat {
    if (length(fibonacci) < 2) {
      next_num <- first  
    } else {
      next_num <- sum(tail(fibonacci, 2))  
    }
    

    if (next_num > 500) {
      break
    }
    
    cat(" ", next_num, sep = "") 
    fibonacci <- c(fibonacci, next_num)  
  }
}

```




```{r}
#4.a
household <- read.table("DataFrame.csv", header = TRUE, sep = ",")
print("Data loaded successfully:")
print(head(household))  
str(household)
```
```{r}
#4.b

male_data <- subset(household, Gender == "M")
female_data <- subset(household, Gender == "F")

male_count <- nrow(male_data)
female_count <- nrow(female_data)

cat("Number of observations in Male:", male_count, "\n")
cat("Number of observations in Female:", female_count, "\n")

```

```{r}
#4.c
gender_counts <- c(male_count, female_count)
names(gender_counts) <- c("M", "F")

barplot(gender_counts,
        main = "Number of Males and Females in Household Data",
        xlab = "Gender",
        ylab = "Count",
        col = c("lightblue", "pink"),
        legend = c("Male", "Female"))

```
```{r}
#5
expenses <- c(Food = 60, Electricity = 10, Savings = 5, Miscellaneous = 25)
expense_labels <- names(expenses)

percent_labels <- paste0(expense_labels, " (", round(expenses / sum(expenses) * 100), "%)")

colors <- c("blue", "lightblue", "navy", "purple")

pie(expenses,
    labels = percent_labels,
    main = "Monthly Expenses of the Dela Cruz Family",
    col = colors)

```
```{r}
#6.a
data(iris)

str(iris)
```
```{r}
#6.b
means <- colMeans(iris[, 1:4])

means

```
```{r}
#6.c
species_counts <- table(iris$Species)

colors <- c("lightgray", "lightblue", "lightgreen")

pie(species_counts,
    main = "Species Distribution in Iris Dataset",
    col = colors,
    labels = paste(names(species_counts), 
                   "(", round(species_counts / sum(species_counts) * 100), "%)", sep = ""))

legend("topright", legend = names(species_counts), fill = colors)

```
```{r}
#6.d
setosa <- subset(iris, Species == "setosa")
versicolor <- subset(iris, Species == "versicolor")
virginica <- subset(iris, Species == "virginica")

cat("Last six rows of Setosa:\n")
tail(setosa)

cat("\nLast six rows of Versicolor:\n")
tail(versicolor)

cat("\nLast six rows of Virginica:\n")
tail(virginica)

```
```{r}
#6.e
iris$Species <- as.factor(iris$Species)

plot(iris$Sepal.Length, iris$Sepal.Width,
     main = "Iris Dataset",
     sub = "Sepal Width and Length",
     xlab = "Sepal Length",
     ylab = "Sepal Width",
     col = c("red", "yellow", "blue")[iris$Species],
     pch = c(16, 17, 18)[iris$Species])

# Add legend
legend("topright", legend = levels(iris$Species), 
       col = c("red", "yellow", "blue"), pch = c(16, 17, 18))

```

#6.f

In the iris dataset, the Sepal.Length versus Sepal.Width scatterplot displays 
separate groups for the setosa, versicolor, and virginica species. Setosa may
be easily distinguished from the other species by its smaller sepal measures,
which often have lower values for both length and width. While Virginica has 
the biggest sepal dimensions, especially in terms of length, Versicolor falls
in the middle, with average sepal diameters. Because each species tends to 
cluster within particular ranges of these measurements, showing natural 
groupings based on sepal size, this separation shows that Sepal.Length and
Sepal.Width are good criteria for categorizing the species.

```{r}
#7.a
library(readxl)  
library(dplyr)  

alexa_data <- read_excel("alexa_file.xlsx")  

print(head(alexa_data))
print(colnames(alexa_data))

if ("Variant" %in% colnames(alexa_data)) { 
    alexa_data$Variant <- gsub("^\\s+|\\s+$", "", alexa_data$Variant)  
    alexa_data$Variant <- gsub("Black Dot", "BlackDot", alexa_data$Variant)
    alexa_data$Variant <- gsub("Black Plus", "BlackPlus", alexa_data$Variant)
    alexa_data$Variant <- gsub("Black Show", "BlackShow", alexa_data$Variant)
    alexa_data$Variant <- gsub("Black Spot", "BlackSpot", alexa_data$Variant)
    alexa_data$Variant <- gsub("White Dot", "WhiteDot", alexa_data$Variant)
    alexa_data$Variant <- gsub("White Plus", "WhitePlus", alexa_data$Variant)
    alexa_data$Variant <- gsub("White Show", "WhiteShow", alexa_data$Variant)
    alexa_data$Variant <- gsub("White Spot", "WhiteSpot", alexa_data$Variant)

    print(head(alexa_data))
} else {
    stop("The specified column 'Variant' does not exist in the dataframe.")
}

```

```{r}
#7.b
if ("Variant" %in% colnames(alexa_data)) {  
    alexa_data$Variant <- gsub("^\\s+|\\s+$", "", alexa_data$Variant)  
    variations_count <- alexa_data %>%
        count(Variant)  
    print(variations_count)

    save(variations_count, file = "variations.RData")  
} else {
    stop("The specified column 'variant' does not exist in the dataframe.")
}
```
```{r}
#7.c
library(ggplot2)

load("variations.RData")  

ggplot(variations_count, aes(x = Variant, y = n, fill = Variant)) +
    geom_bar(stat = "identity") + 
    labs(title = "Count of Variations", 
         x = "Variants", 
         y = "Count") +
    theme_minimal() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
    scale_fill_brewer(palette = "Set3")  

```
```{r}
#7.d
library(gapminder) 
library(dplyr)     
library(ggplot2)   


load("variations.RData") 

black_Variations <- variations_count %>%
    filter(grepl("Black", Variant)) %>%
    summarise(n = sum(n)) %>%
    mutate(color = "Black")  

white_Variations <- variations_count %>%
    filter(grepl("White", Variant)) %>%
    summarise(n = sum(n)) %>%
    mutate(color = "White") 

combined_variations <- rbind(black_Variations, white_Variations)

ggplot(combined_variations, aes(x = color, y = n, fill = color)) +
    geom_bar(stat = "identity", position = "dodge") +  
    labs(title = "Count of Black and White Variations", 
         x = "Variants", 
         y = "Count") +
    theme_minimal() + 
    scale_fill_manual(values = c("Black" = "black", "White" = "gray")) + 
    theme(axis.text.x = element_text(angle = 0, hjust = 0.5))  

```